---
title: "Research Proposal"
author: "Manting Huang"

execute:
  echo : false
  message: false
  warning: false
  
bibliography: references.bib
---

> **Abstract.**  This research project investigates the effects of unanticipated government spending shocks on output, real exchange rate, and consumption, particularly in the context of the Australian economy. Utilizing quarterly data and the SVAR model, the study captures the dynamic relationships among the endogenous variables. The model estimates impulse response functions (IRFs) and forecast error variance decomposition (FEVDs) for each endogenous variable, allowing for the examination of the dynamic responses of variables and quantification of forecast error variances attributable to the shocks. 
>
> **Keywords.** Government spending shock, exchange rate,impulse responses

Please note that the website https://www.rba.gov.au/statistics/tables/ may sometimes be unstable and unable to load. Therefore, you may encounter errors such as "Could not read HTML at https://www.rba.gov.au/statistics/tables/" when using read_rba() in R.


# 1. The question objective, and motivation

## Objective
The research project aims to model the effect of government spending shocks on output, real exchange rate and consumption

## Research question 
How does the unanticipated government spending shock influence output, real exchange rate and consumption in the short run and in the long run?

## Motivation
Most economies have experienced significant shocks due to Covid Pandemic,resulting in a marked increase in government net debt-to-GDP ratios.  For instance,Australian's general government net debt-to-GDP ratio surged from 24.5\% in 2019 to 38.1 in 2021\%  by [@abs2021government], while the United States's ratio rose from 106.1\% to 121.1\% by [@fed2023federal]. 


There has been considerable research on the consequences of government spending shocks on output, consumption, and other crucial macroeconomic factors. Economic theory suggests that positive shocks result in increased output and consumption. However, there has been less emphasis on the economy's external sector, including real exchange rates, imports, and exports. By employing the SVARs model to identify government spending shocks and investigating their connection with these variables, policymakers can enhance the formulation and execution of fiscal and monetary policies in response to unanticipated government spending shocks.

# 2. Data and their properties

## Data 

For this study, quarterly data will be utilized for estimation purposes, as it is more reasonable to assume that government spending can respond within a given period. The primary focus of this research will be on the Australian economy, allowing for an in-depth analysis of its reaction to government spending shocks. It is important to note that the findings may not necessarily be applicable to other countries due to their distinct institutional features and economic attributes. Future research could extend the model to other nations in order to gain a broader understanding of cross-country variations in response to government spending shocks.

Initially, I will load the packages that enable direct data downloads from the Reserve Bank of Australia [@rba2023statistics] and the Federal Reserve Bank of St. Louis [@FRED].

```{r, echo=TRUE}
library(readabs)
library(readrba)
library(tseries)
library(ggplot2)
library(cowplot)
library(dplyr)
library(zoo)
library(tidyverse)
library(fredr)
library(urca)
library(xts)
```



## Variables

### Consumption
The dataset contains information on household final consumption expenditure, which is measured in millions of dollars and recorded quarterly, with adjustments for seasonal variations. The plot, however, reveals that the data is non-stationary, with the mean value changing over time. This is a common characteristic of macroeconomic variables. As a result, I also look at the consumption growth data instead, which measures the year-ended household consumption growth and displays a decline in 2020 due to the COVID-19 pandemic. The mean value of consumption growth is 3.41%, indicating that the growth rate fluctuates around this value and is stationary over time.

In the model, the original consumption variable is log-transformed to reduce the scale and conform to the assumption of normality.



```{r}
consumption <- read_rba('H2')
consumption$Date <- as.Date(consumption$date, format = "%Y-%m-%d")

# Household final consumption expenditure measured in $million

consumption2 <- consumption %>%
  filter(series_id == "GGDPECCVPSH") %>%
  rename(consumption = value)

df <- consumption2 %>% select(Date,consumption)
# Consumption growth
consumption3 <- consumption %>%
  filter(series_id == "GGDPECCVPSHY") %>%
  rename(consumption_growth = value) %>% select(Date,consumption_growth)

df <- merge(df,consumption3,by="Date")

df$log_con = log(df$consumption)


# Plot the data
plot1<- ggplot(data = df, aes(x = Date, y =log_con)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " Household Consumption (log)", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")
plot2<- ggplot(data = df, aes(x = Date, y =consumption_growth)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " Household Consumption Growth", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")

consumption_plot <- plot_grid(plot1, plot2,  ncol = 1, align = "v")
print(consumption_plot)

```
### GDP

The real GDP is measured in million dollars, and the plot of this data suggests that it is non-stationary due to its increasing trend. To address this, I transformed the original variable using the lag() function in R, resulting in the GDP percentage change.

When examining the plot of the GDP percentage change, it appears to be more stationary over time. However, the data fluctuates significantly during the COVID period. In a later section, I also run ACF and PACF tests, which suggest that I should keep the original variable.
```{r}
# GDP
GDP <- read_rba('H1')
GDP$Date <- as.Date(GDP$date, format = "%Y-%m-%d")
# Real GDP

GDP2 <- GDP %>%
  filter(series_id == "GGDPCVGDP") %>%
  rename(GDP = value) %>%
  select(Date,GDP)

df <- merge(df,GDP2, by="Date")

df$log_gdp = log(df$GDP)

# First difference for gdp
df <- df %>% mutate(GDP_percentage = (GDP-lag(GDP))/lag(GDP)*100)

plot2<- ggplot(data = df, aes(x = Date, y =log_gdp)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " GDP (log)", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")
plot3<- ggplot(data = df, aes(x = Date, y =GDP_percentage)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " GDP percentage change", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")

consumption_plot <- plot_grid(plot2, plot3,  ncol = 1, align = "v")
print(consumption_plot)

```
### Exchange rate

The real exchange rate data is sourced from the Reserve Bank of Australia (RBA) and is provided on a quarterly basis. RBA uses the Australian dollar trade-weighted index as a measure of the real exchange rate. This index represents the price of the Australian dollar in terms of a group of foreign currencies, based on their share of trade with Australia.
```{r }


# TWI 

real_exchange_rate <- read_rba("F15")
real_exchange_rate$Date <- as.Date(real_exchange_rate$date, format = "%Y-%m-%d")

real_exchange_rate_twi <-   real_exchange_rate %>%
  filter(series_id == "FRERTWI") %>%
  rename(TWI = value) %>%
  select(Date,TWI)
  
df <- merge(df,real_exchange_rate_twi, by='Date')


plot5<- ggplot(data = df, aes(x = Date, y =TWI)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " Real exchange rate (TWI)", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")

print(plot5)
```






Additionally, I also investigate the nominal exchange rate variable. I download the AUD/USD exchange rate, as the US dollar is widely accepted as an international currency. (Note: The data obtained using the package may not exactly match the description on the website.) However, since the nominal exchange rate may be influenced by various factors and due to data availability constraints, this research will focus exclusively on the real interest rate (TWI).



### Trade balance

The trade balance data is measured as a percentage of output and is recorded on a quarterly basis with seasonal adjustments. Upon examining the plot of this data, it appears to have been stationary before the COVID period. However, since then, the trade balance has been increasing. I will also include this variable in my model as a control variable.


```{r}
Trade_balance <- read_rba("I1")

Trade_balance$Date <- as.Date(Trade_balance$date, format = "%Y-%m-%d")

Trade_balance2 <- Trade_balance   %>%
filter(series_id == "HTBGSCPGDP") %>%
  rename(Trade_balance = value) %>%
  select(Date,Trade_balance)
df <- merge(df,Trade_balance2,by = "Date")

plot6<- ggplot(data = df, aes(x = Date, y =Trade_balance)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " Trade Balance", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")

print(plot6)

```


=======

### Inflation

```{r}
CPI = read_rba("G1")



inflation <- CPI %>%
  filter(series_id == "GCPIAGQP") %>%
  rename(Date = date, Quarterly_inflation = value) %>%
  select(Date, Quarterly_inflation)


df <- merge(df,inflation,by = "Date")
```
### M3

```{r}
M3 <- read_rba("D3")

# Filter, rename, and select columns
M3 <- M3 %>%
  filter(series_id == "DMAM3N") %>%
  rename(Date = date, M3 = value) %>%
  select(Date, M3)

# Create a zoo object
M3.zoo <- zoo(M3$M3, M3$Date)


# Function to convert the index to quarterly and set the last day of each quarter
to_quarter <- function(x) {
  as.yearqtr(format(x, "%Y-%m-%d"), format = "%Y-%m")
}

# Aggregate the monthly data to quarterly
M3_quarterly <- aggregate(M3.zoo, to_quarter, mean)

# Function to convert the yearqtr object to the last day of each quarter
yearqtr_to_quarter_end <- function(x) {
  as.Date(as.yearqtr(x) + 1/4 - 1, frac = 1)
}

# Convert the aggregated quarterly data back to a data frame
M3_quarterly_df <- data.frame(Date = yearqtr_to_quarter_end(index(M3_quarterly)), M3 = coredata(M3_quarterly))

df = merge(df,M3_quarterly_df, by= "Date")

```


### government spending data
Due to package constraints, readrba or readabs are not applicable for government spending data, as the data published about the public sector by ABS is yearly. Instead, I use FRED as my data source. Government spending is increasing due to the upward trend, so I use the lag() function to transform this data.  I will keep the lof transformation of the original variable.
```{r}
# Set your FRED API key

readRenviron("secret.R")
api_key <-Sys.getenv("MY_FRED_API_KEY")
fredr_set_key(api_key)

# Download the Australian government spending data (quarterly data, seasonally adjusted)
gov_spending_australia <- fredr(series_id = "AUSGFCEQDSMEI",observation_start = as.Date("1970-01-01"))

#  Real General Government Final Consumption Expenditure for Australia 
gov_spending_australia <- fredr(series_id = "NCGGRSAXDCAUQ",observation_start = as.Date("1970-01-01"))


gov_spending_australia$log_gov = log(gov_spending_australia$value)

gov_spending_australia <- gov_spending_australia %>%
  mutate(percentage_change = (value - lag(value)) / lag(value) * 100) %>% 
select(date,value,percentage_change,log_gov) %>%
  rename(Date = date)

plot7<- ggplot(data = gov_spending_australia, aes(x = Date, y =log_gov)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " Government Spending", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")

plot8<- ggplot(data = gov_spending_australia, aes(x = Date, y =percentage_change)) +
  geom_line(color = "steelblue", size = 0.8) +
  labs(title = " Government Spending percentage change", x = "Year", y = "Index") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "none")
gov_plot <- plot_grid(plot7, plot8,  ncol = 1, align = "v")
print(gov_plot)

```
I also run the ADF test for this variable, and with a p-value of 0.5, we cannot reject the null hypothesis, indicating that the government spending data is expected to be non-stationary. For completeness, I also plot the ACF (autocorrelation function) and PACF (partial autocorrelation function). As shown in the ACF, there is a slow and gradual decay, suggesting that an autoregressive process might be suitable for the data. The PACF results also support using an AR model because the plot has a shorp cuf-off.
```{r}
# adf.test(gov_spending_australia$log_gov)
# test1 = acf(gov_spending_australia$log_gov)
# test2 = pacf(gov_spending_australia$log_gov)
```

## preliminary Data Analysis

In this section, I will present a table of the ADF test results for the remaining variables for completeness. As stated in the previous section, I will retain the log transformation of the original variable in my research. Additionally, the PACF and ACF tests suggest using an autoregression model.
### ADF test
 

```{r}
# 
# ## Preliminary test
# 
# 
# 
# 
# 
# # Assuming your dataset is called 'data'
# data_without_date <- df[,-1] # Remove date column if it exists
# 
# 
# # Perform the ADF test on each variable and extract the test statistic and p-value
# adf_summary <- map_df(data_without_date, function(x) {
#   adf_test <- ur.df(x, type = "none", lags = 1)
#   adf_test_summary <- summary(adf_test)
#   tibble(
#     Test_statistic = adf_test_summary@teststat[1, 1],
#     Critical_value= adf_test_summary@cval[1, 1]
#   )
# }, .id = "Variable")
# 
# # Add variable names to the summary table
# adf_summary$Variable <- colnames(data_without_date)
# 
# # Display the ADF test summary table
# adf_summary

```

As anticipated, macroeconomic data is typically non-stationary. The only exception in this case is the real exchange rate, which is stationary.

### PACF and ACF test

```{r}
# Define the number of rows and columns for the combined plot
# nrow <- 3
# ncol <- 3
# 
# # Set up the combined plot grid
# par(mfrow = c(nrow, ncol))
# 
# # Run the ACF and PACF functions for each variable and display the plots
# for (i in seq_along(data_without_date)) {
#   variable_name <- colnames(data_without_date)[i]
#   
#   # ACF plot
#   acf(data_without_date[[i]], main = paste("ACF of", variable_name))
# }
# 
# # Reset the graphical parameters to default
# par(mfrow = c(nrow, ncol))
# 
# # PACF plot
# for (i in seq_along(data_without_date)) {
#   variable_name <- colnames(data_without_date)[i]
#   
#   # PACF plot
#   pacf(data_without_date[[i]], main = paste("PACF of", variable_name))
# }

```

# 3. The model 




## Model Specification 
Using same notations in lecture, the model is specified as follows:
$$B_0 y_t = b_0 + B_1 y_{t-1}+ ... + B_p y_{t-p}+u_t$$
Then the empirical model for this study is a structural vector autoregression of the reduced form representation

$$  \begin{bmatrix} \hat{g_t} \\   \hat{c_t} \\ \hat{y_t} \\ \hat{e}_t \\ \hat{i}_t  \\ \hat{m}_t \end{bmatrix} = A(L) \begin{bmatrix} \hat{g}_{t-1} \\   \hat{c}_{t-1} \\ \hat{y}_{t-1} \\ \hat{e}_{t-1} \\\hat{i}_{t-1  }\\ \hat{m}_{t-1} \end{bmatrix} + \varepsilon_t $$
Where:

$g_t$ denotes real government spending,

$c_t$ denotes household final consumption,

$y_t$ denotes real GDP,

$e_t$ denotes the real exchange rate (trade-weighted index),
$i_t$ denotes the inflation,
$m_t$ denotes the M3. 

$A(L)$ is an nxn matrix polynomial in the lag operator L, representing the dynamic relationships among the endogenous variables,

$\varepsilon_t$ is a vector of the structural shocks at time t: $\varepsilon_t = [\varepsilon_{g_t}, \varepsilon_{c_t}, \varepsilon_{y_t}, \varepsilon_{e_t}]^T$.

Assumptions regarding the model's error terms:
$\varepsilon_t |Y_{t-1}  \sim iid(0_N, Σ)$

## Impusle response function and Forecast error variance decomposition

The research project aims to model the effect of government spending shocks on output, real exchange rate, and consumption. The SVAR model specified above captures the dynamic relationships among the endogenous variables. To answer the research questions, I will use the SVAR model to estimate the impulse response functions (IRFs) and forecast error variance decomposition (FEVDs) for each endogenous variable. With IRFs, I will be able to observe the dynamic response of each variable, while the FEVDs will quantify the portions of the forecast error variance of each variable attributable to the shocks this paper aims to identify.


## Relevant economic context
I would expect the IRF for government spending to initially increase in response to the government spending shock and then decrease to the steady state. Consumption and GDP will also increase, but on a smaller scale compared to government spending. I am uncertain about the real exchange rate. Data shows that the real exchange rate will increase upon the shock and continue to rise after the shock.



## References {.unnumbered}